---
title: "Searching best parameter for the in-lab exp"
author: "OSCAR" 
date: "2018 Feb. 12"
output: 
  html_document:
    code_folding: show
    highlight: textmate
    keep_md: no
    number_sections: true
    theme: flatly
    toc: yes
    toc_float:
      collapsed: false
      smooth_scroll: true
      toc_depth: 4
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
if (!require(ggplot2)) {
  install.packages("ggplot2")
  library(ggplot2)
}else{library(ggplot2)}

if (!require(cowplot)) {
  install.packages("cowplot")
  library(cowplot)
}else{library(cowplot)}

if (!require(tidyverse)) {
  install.packages("tidyverse")
  library(tidyverse)
}else{library(tidyverse)}

if (!require(viridis)) {
  install.packages("viridis", dependencies=TRUE, repos='https://ftp.ussg.iu.edu/CRAN/')
  library(viridis)
}else{library(viridis)}
```

```{r, preping/cleaning data, echo=FALSE, warning=FALSE, message=FALSE}
dat <- read.table(file="https://raw.githubusercontent.com/OscarFHC/IGP_LabExp/master/data/IGP_II_20170419_1.csv", sep = ",", header = TRUE, fill = TRUE)
dat[,2:ncol(dat)] <- dat[, 2:ncol(dat)] + 0.01

dat_l <- dat %>% 
  gather(key = trmt, value = den, -hr) %>% 
  mutate(trmt_all = 0)

dat_l[which(grepl("B00", as.character(dat_l[,"trmt"]))==TRUE), "trmt_all"]="B00"
dat_l[which(grepl("C00", as.character(dat_l[,"trmt"]))==TRUE), "trmt_all"]="C00"
dat_l[which(grepl("B02", as.character(dat_l[,"trmt"]))==TRUE), "trmt_all"]="B02"
dat_l[which(grepl("C02", as.character(dat_l[,"trmt"]))==TRUE), "trmt_all"]="C02"
dat_l[which(grepl("B04", as.character(dat_l[,"trmt"]))==TRUE), "trmt_all"]="B04"
dat_l[which(grepl("C04", as.character(dat_l[,"trmt"]))==TRUE), "trmt_all"]="C04"
dat_l[which(grepl("B06", as.character(dat_l[,"trmt"]))==TRUE), "trmt_all"]="B06"
dat_l[which(grepl("C06", as.character(dat_l[,"trmt"]))==TRUE), "trmt_all"]="C06"
dat_l[which(grepl("B08", as.character(dat_l[,"trmt"]))==TRUE), "trmt_all"]="B08"
dat_l[which(grepl("C08", as.character(dat_l[,"trmt"]))==TRUE), "trmt_all"]="C08"
dat_l[which(grepl("B10", as.character(dat_l[,"trmt"]))==TRUE), "trmt_all"]="B10"
dat_l[which(grepl("C10", as.character(dat_l[,"trmt"]))==TRUE), "trmt_all"]="C10"

NonParam_log <- function(den){ # This function log transforms the data when calculatingnon-parametric SE
  set.seed(1032)
  means = replicate(5000, mean(log(sample(den, size=length(den), replace=TRUE))))
  sd(means)
}

NonParam <- function(den){ # This function does not log-transform data before doing permutation
  set.seed(1032)
  means = replicate(5000, mean(sample(den, size=length(den), replace=TRUE)))
  sd(means)
}

dat_sum <- dat_l %>% 
  group_by(trmt_all, hr) %>% 
  summarize(
    rep = length(den),
    avg = mean(den),
    avg_log = mean(log(den)),
    sd = sd(den),
    sd_log = sd(log(den)),
    se = sd/sqrt(rep),
    se_log = sd_log/sqrt(rep),
    se_permu = NonParam(den),
    se_permu_log = NonParam_log(den)
  ) %>%
  ungroup()

tp_start <- 2
tp_end <- length(unique(dat_sum[["hr"]]))-2
timepoint<- unique(dat_sum[["hr"]])[tp_start:tp_end]

i <- 8

B_sum_00 <- dat_sum %>% 
  filter(hr %in% timepoint[i:(i+4)]) %>%
  filter(trmt_all == "B00")

C_sum_00 <- dat_sum %>% 
  filter(hr %in% timepoint[i:(i+4)]) %>%
  filter(trmt_all == "C00")

B_cal <- dat_sum %>%
  filter(hr %in% timepoint[i:(i+4)] & substr(trmt_all, 1, 1) == "B") %>%
  group_by(trmt_all) %>%
  summarize(
    rep = length(avg),
    avg_trmt = mean(avg/mean(B_sum_00[["avg"]])),
    sd_trmt = sd(avg/mean(B_sum_00[["avg"]])),
    se_trmt = sd_trmt/sqrt(rep),
    se_permu = NonParam(avg/mean(B_sum_00[["avg"]]))
  ) %>%
  mutate(alpha = seq(0, 1, by = 0.2))

C_cal <- dat_sum %>%
  filter(hr %in% timepoint[i:(i+4)] & substr(trmt_all, 1, 1) == "C") %>%
  group_by(trmt_all) %>%
  summarize(
    rep = length(avg),
    avg_trmt = mean(avg/mean(C_sum_00[["avg"]])),
    sd_trmt = sd(avg/mean(C_sum_00[["avg"]])),
    se_trmt = sd_trmt/sqrt(rep),
    se_permu = NonParam(avg/mean(C_sum_00[["avg"]]))
  ) %>%
  mutate(alpha = seq(0, 1, by = 0.2))

dat_cal <- rbind(B_cal, C_cal)
```

```{r, loading Bacteria data, echo=FALSE}
Bac_raw = read.table(file="https://raw.githubusercontent.com/OscarFHC/IGP_LabExp/master/data/IGP_II_20170419_bac2.csv",
                     fill=TRUE, header=TRUE, sep=",")

Bac = Bac_raw %>%
  mutate(time = rep(rep(c("T0", "Tend"), each=nrow(Bac_raw)/8), each=4)) %>%
  subset(Gate.Name %in% c("culture population")) %>%
  mutate(trmt = rep(paste0(
                       c(rep(c("00_", "02_", "04_", "06_"), each=3), 
                         rep(("08_"), each=4),
                         rep(c("10_", "ctrl_"), each=5)),
                       c(rep(seq(from=1, to=3), 4),
                         rep(seq(from=1, to=4), 1),
                         rep(seq(from=1, to=5), 2))
                      ), 2)
        ) %>%
  mutate(trmt_all = rep(
                      c(rep(c("Bac00", "Bac02", "Bac04", "Bac06"), each=3), 
                      rep(c("Bac08"), each=4),
                      rep(c("Bac10", "Bacctrl"), each=5)),2)
         ) %>%
  subset(select=c("time", "trmt_all", "trmt", "Gate.Name", "Concentration")) 
  colnames(Bac) = c("time", "trmt_all", "trmt", "pop", "den")

Cons_ctrl = mean(Bac[which(Bac[,"trmt_all"] == "Bacctrl" & Bac[,"time"] == "Tend"), "den"])
Cons_00 = mean(Bac[which(Bac[,"trmt_all"] == "Bac00" & Bac[,"time"] == "Tend"), "den"])
Eqm_00 = mean(Bac[which(Bac[,"trmt_all"] == "Bac00" & Bac[,"time"] == "Tend"), "den"])

Bac_pop <- Bac %>%
  subset(select = -pop) %>%
  spread(key = time, value = den) %>%
  mutate(Consumption = Cons_ctrl - Tend) %>%
  mutate(Change_Cons = Consumption / Cons_00) %>%
  mutate(Change_Eqm = Tend/Eqm_00)

Bac_cal <- Bac_pop %>%
  group_by(trmt_all) %>%
  summarize(
    rep = length(Change_Eqm),
    avg_trmt = mean(Change_Eqm),
    sd_trmt = sd(Change_Eqm),
    se_trmt = sd_trmt/sqrt(rep),
    se_permu = NonParam(Change_Eqm)
  ) %>%
  mutate(alpha = c(seq(0, 1, by = 0.2), 0))

dat_cal <- rbind(dat_cal, Bac_cal)

```

# Hypothesis

According to the model with only type I functional response, I predict that total animal density at the equilibrium ($Z_{tot}$) first increase and then decrease with intra-guild predation ($\alpha$; the proportion of IG prey that is available to the IG predator). Therefore I hypothesize that _resoure density_ at the equilibrium would first decrease and then increase with contact probability ($\alpha$=0). Note that I normalized the resource density at the equilibrium on the condition where contact probability ($\alpha$=0), so that it is comparable across model and experimental results. 

```{r, hypothsis from Type I model, echo=FALSE}
tempDen_tot_H = read.table(file = "https://raw.githubusercontent.com/OscarFHC/IGP_LabExp/master/data/m1_1_m2_150._C1_1_C2_70._TotZ.csv", 
                           header=F, fill=T, sep=",")

tempEigen = read.table(file = "https://raw.githubusercontent.com/OscarFHC/IGP_LabExp/master/data/m1_1_m2_1.5_C1_1_C2_0.7_Eigenval1.csv", header=F, fill=T, sep=",")

ModResults_Ztot_Hump = as.data.frame(matrix(0, nrow(tempDen_tot_H) * ncol(tempDen_tot_H), 4))
ModResults_Ztot_Hump[,1] = rep(seq(0, 0.495, 0.005), nrow(tempDen_tot_H))
ModResults_Ztot_Hump[,2] = rep(seq(0, 1.0, 0.005), each = ncol(tempDen_tot_H))
for (i in 1:nrow(tempDen_tot_H)){
  start = (i-1)*ncol(tempDen_tot_H)+1
  end = i*ncol(tempDen_tot_H)
  ModResults_Ztot_Hump[start:end,3] = as.numeric(tempDen_tot_H[i,])
  ModResults_Ztot_Hump[start:end,4] = as.numeric(tempEigen[i,])
}
colnames(ModResults_Ztot_Hump) = c("ND", "IGP", "Z_tot", "LeadingEigenVal")

ModResults_Ztot_Hump <- ModResults_Ztot_Hump %>%
  group_by(ND) %>%
  mutate(Z_totp = Z_tot/Z_tot[1])

### Reading files for an example showing hump-shaped effects of IGP on total animal density #####
### Creating a figure of hump-shpaped effects of IGP on total animal density #####
Ztot_Hump = 
  ModResults_Ztot_Hump %>% 
  subset(ND %in% c(0, 0.125, 0.25, 0.375)) %>%
  subset(LeadingEigenVal<0) %>%
  arrange(ND) %>%
  ggplot(aes(x = IGP, y = 2 - Z_totp, color = factor(ND))) +
    geom_line(size=1.5) +
    scale_color_viridis(discrete = TRUE,
                        labels=c("100%", "75%", "50%", "25%"),
                        guide=guide_legend(byrow=TRUE, title="% Resource \n Partitioning")
                        )+
    geom_hline(yintercept=1, linetype="solid", size=0.5)+
    lims(x = c(0, 1)) + 
    labs(x = expression(paste("Contact probability (", alpha, ")")),
         y = expression(paste("Resource density at the equilibrium \n(ratio to 0% contact probability)"))
         ) +
    theme_bw() +
    theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          plot.margin = unit(c(0, 0, 0, 0), "cm"),
          legend.position = c(0.95, 0.5),
          legend.justification = c("right", "top"),
          legend.margin = margin(0, 0, 0, 0, "cm"),
          legend.key = element_rect(color = "white", fill = "white"), 
          legend.title = element_text(size = 13),
          legend.text = element_text(size = 13),
          axis.line = element_line(colour = "black"), 
          axis.text = element_text(size=14), 
          axis.title.x = element_text(size = 14, face = "bold", margin=margin(t = 8, r = 0, b = 0, l = 0)),
          axis.title.y = element_text(size = 14, face = "bold", margin=margin(t = 0, r = 0, b = 0, l = 18))
          )
Ztot_Hump
```

To test this hypothesis, I designed an experiment that manipulates the contact probability by installing a $10 \mu m$screen mesh that is permeable to IG prey but not IG predator. With the screen mesh, I manipulated the contact probability of the IG prey by varying the proportion of IG prey that is available to IG predator. 

# Experimental Results

Before confronting the data with model results, I first visualize the population dynamics of IG prey and predator. 

## Population dynamics visualization

First I used the population density of IG predator and IG prey to identify if the system reached the steady state. 

IG predator *Blepharisma*

```{r, pop dynamics for IG predator, echo=FALSE}
dat_sum %>%
  subset(substr(trmt_all, 1, 1) == "B") %>% 
  ggplot(aes(x = hr, y = avg_log, colour = trmt_all)) + 
  geom_errorbar(aes(ymin = avg_log - se_permu_log, ymax = avg_log + se_permu_log), width = 0.1) +
  geom_line() +
  geom_point(size = 4) +
  scale_color_viridis(discrete = TRUE,
                      labels = c("0%", "20%", "40%", "60%", "80%", "100%"),
                      guide = guide_legend(byrow=TRUE, title = expression(paste("Contact \nprobability (", alpha, ")")))
                      ) + 
  labs(x = "Hours",
       y = "Log[density(ind./mL)]",
       title = expression("Average density of IG predator ("* italic("Blespharisma")* ") across time"),
       caption = "error bars represent standard error of the mean") + 
  theme_bw() +
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        plot.margin = unit(c(0, 0, 0, 0), "cm"),
        legend.justification = c("right", "top"),
        legend.margin = margin(0, 0, 0, 0, "cm"),
        legend.key = element_rect(color = "white", fill = "white"), 
        legend.title = element_text(size = 13),
        legend.text = element_text(size = 13),
        axis.line = element_line(colour = "black"), 
        axis.text = element_text(size=14), 
        axis.title.x = element_text(size = 14, face = "bold", margin=margin(t = 8, r = 0, b = 0, l = 0)),
        axis.title.y = element_text(size = 14, face = "bold", margin=margin(t = 0, r = 0, b = 0, l = 18))
        )
```

IG prey *Colpidium*

```{r, pop dynamics for IG prey, echo=FALSE}
dat_sum %>%
  subset(substr(trmt_all, 1, 1) == "C") %>% 
  ggplot(aes(x = hr, y = avg_log, colour = trmt_all)) + 
  geom_errorbar(aes(ymin = avg_log - se_permu_log, ymax = avg_log + se_permu_log), width = 0.1) +
  geom_line() +
  geom_point(size = 4) +
  scale_color_viridis(discrete = TRUE,
                      labels = c("0%", "20%", "40%", "60%", "80%", "100%"),
                      guide = guide_legend(byrow=TRUE, title = expression(paste("Contact \nprobability (", alpha, ")")))
                      ) + 
  labs(x = "Hours",
       y = "Log[density(ind./mL)]",
       title = expression("Average density of IG predator ("* italic("Colpidium")* ") across time"),
       caption = "error bars represent standard error of the mean") + 
  theme_bw() +
  theme(panel.border=element_blank(), 
        panel.grid.major=element_blank(), 
        panel.grid.minor=element_blank(), 
        plot.margin=unit(c(0, 0, 0, 0), "cm"),
        #legend.position = c(0.95, 0.5),
        legend.justification = c("right", "top"),
        legend.margin=margin(0, 0, 0, 0, "cm"),
        legend.key=element_rect(color="white", fill="white"), 
        legend.title=element_text(size=13),
        legend.text=element_text(size=13),
        axis.line = element_line(colour = "black"), 
        axis.text=element_text(size=14), 
        axis.title.x=element_text(size=14, face="bold", margin=margin(t = 8, r = 0, b = 0, l = 0)),
        axis.title.y=element_text(size=14, face="bold", margin=margin(t = 0, r = 0, b = 0, l = 18))
        )
```

For both *Blespharisma* and *Colpidium*, I calculated the mean density for all six treatments (0%, 20%, 40%, 60%, 80% and 100% encounter probability) at the steady state. Because I cannot be 100% sure what are the most representative time points for the steady state, I calculate the mean density of every five consecutive time points. I then moved this five-point time frame gradually forward from hour 34 to hour 535. Finally , I chose the time frame of  hour 320 to 486 since the other time frames are qualitatively the same.

# Confronting model with empirical data

Now I overlay the model predictions onto the IG predator, IG prey, and resource density at the equilibrium measured in empirical experiment. 

## Type I model

```{r, loading Type I model result, echo=FALSE}
dat_mod_I = read.table(file="https://raw.githubusercontent.com/OscarFHC/IGP_LabExp/master/data/Mod_Type1.csv", sep=",", header=TRUE, fill=TRUE)
dat_mod_I = dat_mod_I %>%
  mutate(Res = (R1+R2) / (dat_mod_I[1,"R1"] + dat_mod_I[1,"R2"]),
         Cons = C / (dat_mod_I[1,"C"]),
         Pred = P / (dat_mod_I[1,"P"]))

# Type I R2 for IG predator
B_I_R2 = cbind(B_cal, subset(dat_mod_I, alpha %in% c(0, 0.2, 0.4, 0.6, 0.8, 1.0)))
SST_I_pred = sum((B_I_R2[,"avg_trmt"] - mean(B_I_R2[,"avg_trmt"]))^2)
SSR_I_pred = sum((B_I_R2[,"avg_trmt"] - B_I_R2[,"Pred"])^2)
R2_I_pred = ifelse((1-(SSR_I_pred/SST_I_pred)) > 0, 1-(SSR_I_pred/SST_I_pred), 0)

# Type I R2 for IG prey
C_I_R2 = cbind(C_cal, subset(dat_mod_I, alpha %in% c(0, 0.2, 0.4, 0.6, 0.8, 1.0)))
SST_I_prey = sum((C_I_R2[,"avg_trmt"] - mean(C_I_R2[,"avg_trmt"]))^2)
SSR_I_prey = sum((C_I_R2[,"avg_trmt"] - C_I_R2[,"Cons"])^2)
R2_I_prey = ifelse((1-(SSR_I_prey/SST_I_prey)) > 0, (1-(SSR_I_prey/SST_I_prey)), 0)

# Type I R2 for Bacteria
Bac_I_R2 = cbind(Bac_cal[-nrow(Bac_cal),], subset(dat_mod_I, alpha %in% c(0, 0.2, 0.4, 0.6, 0.8, 1.0)))
SST_I_bac = sum((Bac_I_R2[,"avg_trmt"] - mean(Bac_I_R2[,"avg_trmt"]))^2)
SSR_I_bac = sum((Bac_I_R2[,"avg_trmt"] - Bac_I_R2[,"Res"])^2)
R2_I_bac = ifelse((1-(SSR_I_bac/SST_I_bac)) > 0, (1-(SSR_I_bac/SST_I_bac)), 0)
```

__IG predator (*Blepharisma*)__

```{r, Pred vs Type I model, echo=FALSE}
ggplot() +
  geom_point(data = B_I_R2, aes(x = alpha, y = avg_trmt), size = 2) +  
  geom_errorbar(data = B_I_R2, aes(x = alpha, ymin = avg_trmt - se_permu, ymax = avg_trmt + se_permu), width = 0.1) + 
  geom_line(data = dat_mod_I, aes(x = alpha, y = Pred), size = 1, color = "blue") + 
  labs(x=expression(paste("Contact probability (", alpha, ")")),
       y=expression(paste("IG predator density at the equilibrium \n(ratio to 0% contact probability)")),
       caption=expression("Error bars represent standard error of the mean \nDashed line represents model results"))+ 
theme_bw() +
theme(panel.border = element_blank(), 
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      plot.margin = unit(c(0, 0, 0, 0), "cm"),
      plot.caption = element_text(size = 10, margin = margin(t = 12, r = 0, b = 0, l = 0)),
      axis.line = element_line(colour = "black"), 
      axis.text = element_text(size = 14), 
      axis.title.x = element_text(size = 14, face = "bold", margin = margin(t = 8, r = 0, b = 0, l = 0)),
      axis.title.y = element_text(size = 14, face = "bold", margin = margin(t = 0, r = 0, b = 0, l = 18))
      )
```

For the IG predator, the model with type I functional response explains `r round(R2_I_pred, 4)*100`% of the empirical data. 
__IG prey (*Colpidium*)__

```{r, Pry vs Type I model, echo=FALSE}
ggplot() +
  geom_point(data = C_I_R2, aes(x = alpha, y = avg_trmt), size = 2) +  
  geom_errorbar(data = C_I_R2, aes(x = alpha, ymin = avg_trmt - se_permu, ymax = avg_trmt + se_permu), width = 0.1) + 
  geom_line(data = dat_mod_I, aes(x = alpha, y = Cons), size = 1, color = "blue") + 
  labs(x=expression(paste("Contact probability (", alpha, ")")),
       y=expression(paste("IG prey density at the equilibrium \n(ratio to 0% contact probability)")),
       caption=expression("Error bars represent standard error of the mean \nDashed line represents model results"))+ 
  theme_bw() +
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        plot.margin = unit(c(0, 0, 0, 0), "cm"),
        plot.caption = element_text(size = 10, margin = margin(t = 12, r = 0, b = 0, l = 0)),
        axis.line = element_line(colour = "black"), 
        axis.text = element_text(size = 14), 
        axis.title.x = element_text(size = 14, face = "bold", margin = margin(t = 8, r = 0, b = 0, l = 0)),
        axis.title.y = element_text(size = 14, face = "bold", margin = margin(t = 0, r = 0, b = 0, l = 18))
        )
```

For the IG prey, the model with type I functional response explains `r round(R2_I_prey, 4)*100`% of the empirical data.

__Bacteria (resource)__

```{r, Bac vs Type I model, echo=FALSE}
ggplot() +
  geom_point(data = Bac_I_R2, aes(x = alpha, y = avg_trmt), size = 2) +  
  geom_errorbar(data = Bac_I_R2, aes(x = alpha, ymin = avg_trmt - se_permu, ymax = avg_trmt + se_permu), width = 0.1) + 
  geom_line(dat = dat_mod_I, aes(x = alpha, y = Res), size = 1, color = "blue") + 
  labs(x=expression(paste("Contact probability (", alpha, ")")),
       y=expression(paste("Bacteria density at the equilibrium \n(ratio to 0% contact probability)")),
       caption=expression("Error bars represent standard error of the mean \nDashed line represents model results"))+ 
theme_bw() +
theme(panel.border = element_blank(), 
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      plot.margin = unit(c(0, 0, 0, 0), "cm"),
      plot.caption = element_text(size = 10, margin = margin(t = 12, r = 0, b = 0, l = 0)),
      axis.line = element_line(colour = "black"), 
      axis.text = element_text(size = 14), 
      axis.title.x = element_text(size = 14, face = "bold", margin = margin(t = 8, r = 0, b = 0, l = 0)),
      axis.title.y = element_text(size = 14, face = "bold", margin = margin(t = 0, r = 0, b = 0, l = 18))
      )
```

For the Bacteria, the model with type I functional response explains `r round(R2_I_bac, 4)*100`% of the empirical data.

The empirical data only qualitatively confirm the IG prey and resource density at the equilibrium and the R^2 is fairly low. The IG predator density at the equilibrium can not be described by the model with just type I functional response. 

## Type II model

From the literature[@Laybourn1975; @Fenchel1980], I found that both IG predator and IG prey have type II functional response when consuming bacteria. However, attack rate and handling time of IG predator are not specifically estimated. I therefore extracted the attack rate and handling time from [@Laybourn1975] for IG prey and used similar value for IG predator to build another more mechanistic model. 

In addition, I did another independent experiment to estimate the attack rate and handling time for IG predator when consuming IG prey. 

```{r, Type2 functional response of IG predator, echo=FALSE}
BC_Type2=read.table(file="https://raw.githubusercontent.com/OscarFHC/IGP_LabExp/master/data/BonC_Type2_20171112.csv", sep=",",header=T,fill=T)

BC_nls = BC_Type2 %>% 
  group_by(Trmt, rep) %>%
  summarize(
    InitRatio = mean(T0),
    End_W = mean(T4_W),
    End_WO = mean(T4_WO),
    Cons = (End_WO-End_W)
    )

Type2_mod = nls(Cons~(s2*InitRatio)/(1+(h2*s2*InitRatio)), start=list(s2=0.4, h2=0.34), data=BC_nls)

BC_nls %>%
  ggplot(aes(x = InitRatio)) +
  geom_point(aes(y = Cons), size = 3) + 
  geom_line(aes(y = predict(Type2_mod)), linetype = 2)+
  labs(x = "IG prey density per IG predator (ind./ml)", 
       y = expression("Number of IG prey consumed (ind./ " * mL%.%day * ")")
       )+ 
  theme_bw() +
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        plot.margin = unit(c(0, 0, 0, 0), "cm"),
        axis.line = element_line(colour = "black"), 
        axis.text = element_text(size = 14), 
        axis.title.x = element_text(size = 14, face = "bold", margin = margin(t = 8, r = 0, b = 0, l = 0)),
        axis.title.y = element_text(size = 14, face = "bold", margin = margin(t = 0, r = 6, b = 12, l = 12))
        )
```

The attack rate and the handing time for IG predator when consuming IG prey are estimated to be `r round(summary(Type2_mod)$coefficients[1,1], 4)` and `r round(summary(Type2_mod)$coefficients[2,1], 4)` respectively. This means the maximum consumption per day should be `r round(1/summary(Type2_mod)$coefficients[2,1], 4)` (IG prey/IG predator) and the half saturation density of IG prey is `r round(1/(summary(Type2_mod)$coefficients[2,1]*summary(Type2_mod)$coefficients[1,1]), 4)`.

However, I cannot find empirical parameter values for the assimilation efficienty of IG prey and the degree of resource partitioning. I search for a range of parameter space to find a combination of IG prey assimilation efficienty and degree of resource partitioning that renders the lowest residual sum of squared. 

```{r, plot estimating parameters for Type II functional response, echo=FALSE}
mod.lst <- expand.grid(e2 = seq(0, 0.5, 0.01), partition = seq(0, 0.15, 0.01))

for (i in 1:nrow(mod.lst)){
  tempdat <- read.table(file = paste0("https://raw.githubusercontent.com/OscarFHC/IGP_LabExp/master/Analysis/Type2ModResults/Type2_e2_", mod.lst[i, "e2"]*10, "_p_", mod.lst[i,"partition"]*100, "..csv"), sep=",", header=FALSE)
  colnames(tempdat) <- c("alpha", "P1", "P2", "C", "B")

  PredSSR <- cbind(B_cal, subset(tempdat, alpha %in% c(0, 0.2, 0.4, 0.6, 0.8, 1.0)))
  PredSSR[,"B"] <- PredSSR[,"B"]/PredSSR[1,"B"]
  mod.lst[i, "RSS_Pred"] <- sum((PredSSR[,"avg_trmt"] - PredSSR[,"B"])^2)/
                            sum((PredSSR[,"avg_trmt"] - mean(PredSSR[,"avg_trmt"]))^2)
  
  PreySSR <- cbind(C_cal, subset(tempdat, alpha %in% c(0, 0.2, 0.4, 0.6, 0.8, 1.0)))
  PreySSR[,"C"] <- PreySSR[,"C"]/PreySSR[1,"C"]
  mod.lst[i, "RSS_Prey"] <- sum((PreySSR[,"avg_trmt"] - PreySSR[,"C"])^2)/
                            sum((PreySSR[,"avg_trmt"] - mean(PreySSR[,"avg_trmt"]))^2)

  ResSSR <- cbind(Bac_cal[-nrow(Bac_cal),], subset(tempdat, alpha %in% c(0, 0.2, 0.4, 0.6, 0.8, 1.0)))
  ResSSR[,"Bac"] <- (ResSSR[,"P1"] + ResSSR[,"P2"]) / (ResSSR[1,"P1"] + ResSSR[1,"P2"])
  mod.lst[i, "RSS_Bac"] <- sum((ResSSR[,"avg_trmt"] - ResSSR[,"Bac"])^2)/
                           sum((ResSSR[,"avg_trmt"] - mean(ResSSR[,"avg_trmt"]))^2)
}

mod.lst <- mod.lst %>%
  mutate(RSS_tot = RSS_Bac + RSS_Prey + RSS_Pred) 

mod.lst %>%
  subset(RSS_Pred<1 & RSS_Prey<1 & RSS_Bac<1) %>%
  ggplot(aes(x = e2, y = 1-partition, fill = RSS_tot)) +
  geom_tile(color = "white", size = 0.1) +
  scale_fill_viridis() + 
  #coord_equal() + 
  labs(x = "Assimilation efficiency of IG prey", 
       y = "Degree of resource partitioning", 
       fill = "Normalized total \nresidual sum of squared") +
  theme_bw() +
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        plot.margin = unit(c(0, 0, 0, 0), "cm"),
        legend.title = element_text(size = 13),
        legend.text = element_text(size = 13),
        legend.justification = c("right", "center"),
        legend.margin = margin(0, 0, 0, 0, "cm"),
        legend.key = element_rect(color = "white", fill = "white"),
        axis.line = element_line(colour = "black"), 
        axis.text = element_text(size=14), 
        axis.title.x = element_text(size = 14, face = "bold", margin=margin(t = 8, r = 0, b = 0, l = 0)),
        axis.title.y = element_text(size = 14, face = "bold", margin=margin(t = 0, r = 6, b = 0, l = 18))
        )
```

From the above figure, the when the assimilation efficienty of IG prey is `r subset(mod.lst, RSS_tot == min(mod.lst[,"RSS_tot"]))[["e2"]]` and the degree of resource partitioning is `r subset(mod.lst, RSS_tot == min(mod.lst[,"RSS_tot"]))[["partition"]]`, the model explain the highest variation of the data. 

With these parameter values, I overlay the model prediction and data again. 

```{r, loading Type II model result, echo=FALSE}
dat_mod_II = read.table(file="https://raw.githubusercontent.com/OscarFHC/IGP_LabExp/master/data/Mod_Type2_e2_4.5_p_5..csv", sep=",", header=FALSE, fill=TRUE)
colnames(dat_mod_II) <- c("alpha", "R1", "R2", "Cons", "Pred")
dat_mod_II = dat_mod_II %>%
  mutate(Res = (R1+R2) / (dat_mod_II[1,"R1"] + dat_mod_II[1,"R2"]),
         Cons = Cons / (dat_mod_II[1,"Cons"]),
         Pred = Pred / (dat_mod_II[1,"Pred"]))

# Type I R2 for IG predator
B_II_R2 = cbind(B_cal, subset(dat_mod_II, alpha %in% c(0, 0.2, 0.4, 0.6, 0.8, 1.0)))
SST_II_pred = sum((B_II_R2[,"avg_trmt"] - mean(B_II_R2[,"avg_trmt"]))^2)
SSR_II_pred = sum((B_II_R2[,"avg_trmt"] - B_II_R2[,"Pred"])^2)
R2_II_pred = ifelse((1-(SSR_II_pred/SST_II_pred)) > 0, 1-(SSR_II_pred/SST_II_pred), 0)

# Type I R2 for IG prey
C_II_R2 = cbind(C_cal, subset(dat_mod_II, alpha %in% c(0, 0.2, 0.4, 0.6, 0.8, 1.0)))
SST_II_prey = sum((C_II_R2[,"avg_trmt"] - mean(C_II_R2[,"avg_trmt"]))^2)
SSR_II_prey = sum((C_II_R2[,"avg_trmt"] - C_II_R2[,"Cons"])^2)
R2_II_prey = ifelse((1-(SSR_II_prey/SST_II_prey)) > 0, (1-(SSR_II_prey/SST_II_prey)), 0)

# Type I R2 for Bacteria
Bac_II_R2 = cbind(Bac_cal[-nrow(Bac_cal),], subset(dat_mod_II, alpha %in% c(0, 0.2, 0.4, 0.6, 0.8, 1.0)))
SST_II_bac = sum((Bac_II_R2[,"avg_trmt"] - mean(Bac_II_R2[,"avg_trmt"]))^2)
SSR_II_bac = sum((Bac_II_R2[,"avg_trmt"] - Bac_II_R2[,"Res"])^2)
R2_II_bac = ifelse((1-(SSR_II_bac/SST_II_bac)) > 0, (1-(SSR_II_bac/SST_II_bac)), 0)
```

__IG predator (*Blepharisma*)__

```{r, Pred vs Type II model, echo=FALSE}
ggplot() +
  geom_point(data = B_II_R2, aes(x = alpha, y = avg_trmt), size = 2) +  
  geom_errorbar(data = B_II_R2, aes(x = alpha, ymin = avg_trmt - se_permu, ymax = avg_trmt + se_permu), width = 0.1) + 
  geom_line(data = dat_mod_II, aes(x = alpha, y = Pred), size = 1, color = "blue") + 
  labs(x = expression(paste("Contact probability (", alpha, ")")),
       y = expression(paste("IG predator density at the equilibrium \n(ratio to 0% contact probability)")),
       caption = expression("Error bars represent standard error of the mean \nDashed line represents model results"))+ 
  theme_bw() +
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        plot.margin = unit(c(0, 0, 0, 0), "cm"),
        plot.caption = element_text(size = 10, margin = margin(t = 12, r = 0, b = 0, l = 0)),
        axis.line = element_line(colour = "black"), 
        axis.text = element_text(size = 14), 
        axis.title.x = element_text(size = 14, face = "bold", margin = margin(t = 8, r = 0, b = 0, l = 0)),
        axis.title.y = element_text(size = 14, face = "bold", margin = margin(t = 0, r = 0, b = 0, l = 18))
        )
```

For the IG predator, the model with type II functional response explains `r round(R2_II_pred, 4)*100`% of the empirical data. 
__IG prey (*Colpidium*)__

```{r, Prey vs Type II model, echo=FALSE}
ggplot() +
  geom_point(data = C_II_R2, aes(x = alpha, y = avg_trmt), size = 2) +  
  geom_errorbar(data = C_II_R2, aes(x = alpha, ymin = avg_trmt - se_permu, ymax = avg_trmt + se_permu), width = 0.1) + 
  geom_line(data = dat_mod_II, aes(x = alpha, y = Cons), size = 1, color = "blue") + 
  labs(x = expression(paste("Contact probability (", alpha, ")")),
       y = expression(paste("IG prey density at the equilibrium \n(ratio to 0% contact probability)")),
       caption = expression("Error bars represent standard error of the mean \nDashed line represents model results"))+ 
  theme_bw() +
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        plot.margin = unit(c(0, 0, 0, 0), "cm"),
        plot.caption = element_text(size = 10, margin = margin(t = 12, r = 0, b = 0, l = 0)),
        axis.line = element_line(colour = "black"), 
        axis.text = element_text(size = 14), 
        axis.title.x = element_text(size = 14, face = "bold", margin = margin(t = 8, r = 0, b = 0, l = 0)),
        axis.title.y = element_text(size = 14, face = "bold", margin = margin(t = 0, r = 0, b = 0, l = 18))
        )
```

For the IG prey, the model with type II functional response explains `r round(R2_II_prey, 4)*100`% of the empirical data.

__Bacteria (resource)__

```{r, Bac vs Type II model, echo=FALSE}
ggplot() +
  geom_point(data = Bac_II_R2, aes(x = alpha, y = avg_trmt), size = 2) +  
  geom_errorbar(data = Bac_II_R2, aes(x = alpha, ymin = avg_trmt - se_permu, ymax = avg_trmt + se_permu), width = 0.1) + 
  geom_line(data = dat_mod_II, aes(x = alpha, y = Res), size = 1, color = "blue") + 
  labs(x = expression(paste("Contact probability (", alpha, ")")),
       y = expression(paste("Bacteria density at the equilibrium \n(ratio to 0% contact probability)")),
       caption = expression("Error bars represent standard error of the mean \nDashed line represents model results"))+ 
  theme_bw() +
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        plot.margin = unit(c(0, 0, 0, 0), "cm"),
        plot.caption = element_text(size = 10, margin = margin(t = 12, r = 0, b = 0, l = 0)),
        axis.line = element_line(colour = "black"), 
        axis.text = element_text(size = 14), 
        axis.title.x = element_text(size = 14, face = "bold", margin = margin(t = 8, r = 0, b = 0, l = 0)),
        axis.title.y = element_text(size = 14, face = "bold", margin = margin(t = 0, r = 0, b = 0, l = 18))
        )
```

For the Bacteria, the model with type II functional response explains `r round(R2_I_bac, 4)*100`% of the empirical data.

# Explanation

```{r, calculation predator prey density ratio, echo=FALSE}
mean_den <- dat_sum %>%
  filter(hr %in% c(320, 368, 414, 468, 486)) %>%
  group_by(trmt_all) %>%
  summarize(
    Den_mean = mean(avg)
  )

PPratio_04 <- (mean_den[which(mean_den[["trmt_all"]] == "C04"),"Den_mean"] * 0.4) / 
              mean_den[which(mean_den[["trmt_all"]] == "B04"),"Den_mean"]

PPratio_08 <- (mean_den[which(mean_den[["trmt_all"]] == "C08"),"Den_mean"] * 0.8) / 
              mean_den[which(mean_den[["trmt_all"]] == "B08"),"Den_mean"]
```

When contact probability is low ($\alpha$ = 20%), only small proportion of IG prey is being consumed by IG predator. The main interaction between IG prey and IG predator is still competition, so bacteria consumption is not different from the treatment where IGP is not allowed to occur ($\alpha$= 0%). When contact probability gradually increases to 40-60%, the IG predator reached higher density but the IG prey is not being completely suppressed as the IG predator is about to be satiated by IG prey when $\alpha$ reaches 40% (the IG prey to IG predator ratio is `r round(PPratio_04, 4)` and the half saturation density of IG prey is `r round(1/(summary(Type2_mod)$coefficients[2,1]*summary(Type2_mod)$coefficients[1,1]), 4)`). Under this treatment, bacteria density is being suppressed to lowest level by high IG predator density and intermediate IG prey density. When contact probability increase further higher to over 80%, IG predator is satiated by the IG prey (the IG prey to IG predator ratio is over `r round(PPratio_08, 4)`). The IG prey density therefore does not decrease further. The reduction of IG predator could be due to competition incurred by the IG prey. The lowest level of IG prey density and satiated IG predator result in highest resource density at the equilibrium.